04_analysis

### 2023-06-29 04_analysis.Rmd
Update 04_analysis.Rmd

print(ggplot(data = pivot1, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 3, color = "black") +
  theme_classic() +
  ggtitle(paste0("Distribution of third octave level intensities: Drift ", substr(ft,10,11))) +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.7,0.8)) +
  facet_grid(~SL))
}

@@

ggplot(data = pivot3, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 3, color = "black") +
  theme_classic() +
  ggtitle("Distribution of third octave level intensities: All Drifts Combined") +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.7,0.8)) +
  facet_grid(~SL)

@@

@@

```{r}
model3 <- gam(BWpresence ~ depth + TOL_125 + dist2shore + mldTemp, 
              data = allDrifts, family = binomial)
plot(model3)

@@

### Plot maps for Powerpoint images

Set Parameters
```{r}
theme_set(theme_bw())
tmap_mode("plot")
```

Create sf objects and basemaps for maps (not needed for environmental variables)
```{r}
whales_8_10sf <- Wsf %>%                            
dplyr::filter(station == "8" | station == "10")

trackssf <- st_as_sf(tracks, coords = c("Longitude", "Latitude"), crs=4326)
trk_TOL_08 <- st_as_sf(tracks_SS_08, coords = c("Longitude", "Latitude"), crs=4326)
trk_TOL_10 <- st_as_sf(tracks_SS_10, coords = c("Longitude", "Latitude"), crs=4326)
trackBase <- get_tiles(whales_8_10sf, provider = "Esri.WorldShadedRelief")
allTrackbase <- get_tiles(trackssf, provider = "Stamen.TonerLite", zoom = 5)
```

Map of whole study area
```{r}
# Establish bounding box
bbox <- st_bbox(c(xmin = -110, 
            xmax = -128, 
            ymax = 44, 
            ymin = 30), crs = st_crs(4326))

#Establish basemap
allTrackbase <- get_tiles(bbox, provider = "Esri.WorldGrayCanvas", zoom = 5)

# Establish inset map
us_region <- st_bbox(c(xmin = -90,
            xmax = -120, 
            ymax = 48, 
            ymin = 25), crs = st_crs(4326))
usa <- get_tiles(us_region, provider = "Esri.WorldGrayCanvas", zoom = 4)

sg <- bb_poly(bb(tracks_sf))
inset <- tm_shape(usa) + tm_rgb() + tm_shape(sg) + tm_borders(col = "red")

#Map all tracks
main <- tm_shape(allTrackbase) + tm_rgb() +
  tm_shape(trackssf) +
  tm_grid(alpha = 0.6, col = "grey") +
  tm_dots(col = "darkblue", size = 0.001, legend.show = TRUE) +
  tm_scale_bar(width = 0.3, text.size = 3, position = c(0.2, 0.03)) +
  tm_compass(type = "arrow", position = c(0.05, 0.05), color.light = "white") +
  tm_add_legend('symbol',
                type = "line",
                col = "darkblue",
                labels = "Drifting buoy tracks",
                size = 1,
                lwd = 2)

vp <- viewport(x=0.66, y=0.8, width = 0.26, height=0.4)
main
print(inset,vp=vp)

# Map all tracks - ggplot
basemap2 <- as.data.frame(allTrackbase, xy = TRUE)
#ggplot() +
#  geom_tile(data = basemap, aes(x=x, y=y, fill = "basemap")) +
#  geom_sf(data = trackssf, mapping = aes(), color = "red")

# Map tracks 08 and 10 only
tm_shape(trackBase) + tm_rgb() + 
  tm_shape(trk_TOL_08) + tm_dots(col = "red", 
                                 legend.show = TRUE) +
  tm_shape(trk_TOL_10) + tm_dots(col = "green")

# Bounding box of drift tracks on map
lon2 <- -129.04666
lon1 <- -115.78247
lat1 <- 28.21790
lat2 <- 45.09226

basemap3 <- ne_countries(scale = "medium", type = "map_units", returnclass = "sf")

# Drift tracks
ggplot(data = basemap3) +
  geom_sf(fill = "antiquewhite") +
  coord_sf(xlim = c(lon2 - 1, lon1 + 1), ylim = c(lat1 - 1, lat2 + 1), expand = FALSE) +
  theme(panel.grid.major = element_blank()) +
  geom_path(data = tracks_sf, mapping = aes(x=long,y=lat, group = factor(station)), color = "deeppink") +
  theme(panel.background = element_rect(fill = "aliceblue"), 
        axis.text = element_blank(), axis.ticks = element_blank(), 
        axis.title = element_blank())
```

Map of soundscapes
```{r}
# CCES_08 + CCES_10 Broadband
tm_shape(trackBase) + tm_rgb() + 
  tm_shape(trk_TOL_08) + 
  tm_symbols(col = "BB_20-24000", 
             border.lwd = NA, 
             size = 0.4, 
             sizes.legend = 0.7, 
             n = 10) +
  tm_shape(trk_TOL_10) + 
  tm_symbols(col = "BB_20-24000",
             border.lwd = NA, 
             size = 0.4,
             n = 10) +
  tm_layout(legend.width = 0.5, 
            legend.position = c("right","top"),
            legend.bg.color = "lightblue")

#CCES_08 Third Octave Level
tm_shape(trackBase) + tm_rgb() + 
  tm_shape(trk_TOL_08) + 
  tm_symbols(col = "TOL_63", 
             border.lwd = NA, 
             size = 0.5, 
             n = 10) + 
  tm_layout(legend.width = 0.5, 
            legend.position = c("right","top")) + 
  tm_legend()

ggplot() + 
  geom_sf(data = trk_TOL_08, aes(color = TOL_125, size = TOL_125), show.legend = FALSE) +
  geom_sf(data = trk_TOL_10, aes(color = TOL_125, size = TOL_125)) +
  scale_x_continuous(limits = c(-130,-122)) + 
  scale_y_continuous(limits = c(35,39)) +
  scale_color_gradient(low = "navy", high = "yellow")+
  scale_size(range = c(1,12)) +
  annotation_scale(location = "br", width_hint = 0.2, plot_unit = "km") +
  theme(panel.border = element_rect(fill = NA),
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        panel.background = element_blank(),
        legend.position = "none")
```

Map of whale detections
```{r}
# Whale detections
ggplot() +
  geom_point(data = whales8, mapping = aes(x=long, y=lat), shape = 21, color = "black", fill = "green4", size = 4) +
  geom_point(data = whales10, mapping = aes(x=long, y=lat), shape = 21, color = "black", fill = "green4", size = 4) +
  scale_x_continuous(limits = c(-130,-122)) + 
  scale_y_continuous(limits = c(35,39)) +
  annotation_scale(location = "br", width_hint = 0.2, plot_unit = "km") +
  theme(panel.border = element_rect(fill = NA),
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        panel.background = element_blank()
        )
#ggsave("Wdetect.png")

Map of bathymetry
```{r}
# Bathymetry
bathym <- ggplot() +
    geom_raster(data = bathy_df, 
              mapping = aes(x=x, y=y,
                            fill = `gebco_bathy`)) +
  annotation_scale(location = "bl", width_hint = 0.4) +
  theme(axis.text = element_blank(), 
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA))

# Alternative Bathymetry (code from Ocean Tracking Network #https://www.youtube.com/watch?v=xT2F-Y9bKLk)
bgo <- getNOAA.bathy(lon1 = -122, lon2 = -130,
                     lat1 = 35, lat2= 39, resolution = 1)  # plot bounding box around drift 8 and 10
autoplot(bgo, geom = c("raster", "contour"), coast = TRUE, show.legend = FALSE) + 
  scale_fill_etopo() + #creates ggplot object
  annotation_scale(location = "br", width_hint = 0.2, plot_unit = "km") +
  theme(axis.text = element_blank(), 
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        panel.background = element_blank())

# Soundscape + bathymetry
whales08_sf <- Wsf %>%                            
  dplyr::filter(station == "8")

ggplot() +
  geom_raster(data = bathy_df, 
              mapping = aes(x=x, y=y,
                            fill = `gebco_bathy`), show.legend = FALSE) +
  #geom_contour(data = bathy_df, mapping = aes(x=x, y=y, z=gebco_bathy, linewidth = 0.1), color = "blue", linewidth = 0.01,) +
  geom_sf(data = trk_TOL_08, aes(color = TOL_125, size = TOL_125), show.legend = FALSE) +
  geom_sf(data = whales08_sf, aes(), color = "black") +
  annotation_scale(location = "br", width_hint = 0.2, plot_unit = "km") +
  scale_color_gradient(low = "beige", high = "hotpink") +
  scale_fill_etopo() +
  theme(axis.text = element_blank(), 
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA))
#ggsave("soundscape.png")
```

Map of AIS
```{r}
# Plot + crop to drift area
#basemap <- ne_countries(scale = "medium", returnclass = "sf")
#ggplot() + 
#  geom_sf(aes(), basemap) +
#  geom_path(data = AIS_sf, mapping = aes(x=LON,y=LAT, group = VesselName), #show.legend = F) + 
  #geom_point(data = AIS_sf, mapping = aes(x=LON,y=LAT, group = VesselName)) +
#  scale_x_continuous(limits = c(min(drift$long-0.7), max(drift$long + 0.7))) + 
#  scale_y_continuous(limits = c(min(drift$lat-0.7), max(drift$lat + 0.7)))

# AIS data
ggplot() + 
  geom_path(data = AIS_sf, mapping = aes(x=LON,y=LAT, group = VesselName, linewidth = 1), show.legend = F) +
  scale_x_continuous(limits = c(-131,-121)) + 
  scale_y_continuous(limits = c(34,40)) +
  theme(axis.text = element_blank(), 
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA))
#ggsave("AIS.png")
```

Map of Sea Surface Temp with rerddap, Drift 8 only
```{r}
# Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01°, 2002-present, Daily
SSTInfo <- info('jplMURSST41') 
murSST <- griddap(SSTInfo, latitude = c(35, 39), longitude = c(-122, -130), 
                  time = c("2018-08-16", "2018-10-01"), 
                  fields = 'analysed_sst') # drift 8 dates only
SST <- murSST$data

# Raster of SST
ggplot() + 
  geom_tile(data = SST, aes(x = longitude, y = latitude, fill = analysed_sst)) +
  scale_fill_gradient(low = "darkblue", high = "yellow") +
  annotation_scale(location = "br", width_hint = 0.2, plot_unit = "km") +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(fill = NA),
        legend.position = "none")
```

Plot distance to continental slope
```{r}
ggplot() +
  geom_raster(data = bathy_df, 
              mapping = aes(x=x, y=y,
                            fill = `gebco_bathy`)) +
  geom_sf(data = tracks_SS_08, 
          mapping = aes(col = dist_slope)) +
  geom_sf(data = cont_slope) + scale_fill_continuous(high = "lightgreen", low = "blue")

# PLOT distance by TOL with whales
ggplot(data = tracks_SS_08, aes(x = dist_slope, y = `TOL_125`)) + 
  geom_point() + 
  geom_smooth(color = "green") +
  theme_classic()
```

Another map - need to decide if important to keep
```{r}
# Plot
oceanBase <- get_tiles(tracks_SS_08, provider="Esri.NatGeoWorldMap", crop = TRUE)
tm_shape(oceanBase) + tm_rgb() +
  tm_shape(tracks_SS_08) +  tm_symbols(col="species", size = 1) +
  tm_layout(legend.width = 0.5, 
            legend.position = c("right","top"))

#Raster base map
oceanBase <- get_tiles(tracks_SS_10, provider="Esri.NatGeoWorldMap")
tm_shape(bathy) + tm_raster() +
  tm_shape(tracks_SS_10) + 
  tm_symbols(style = "jenks") +
  tm_layout(legend.width = 0.5)
```

```{r}
# Plots for coastline and continental slope. Do I keep?
sf08 <- st_as_sf(tracks_SS_08, coords = c("Longitude", "Latitude"), crs=4326, remove=F)
sf10 <- st_as_sf(tracks_SS_10, coords = c("Longitude", "Latitude"), crs=4326, remove=F)
trkraster08 <- rasterize(vect(sf08), agg)
trkraster10 <- rasterize(vect(sf10), agg)

bathy_df <- as.data.frame(bathy, xy = TRUE)

# Map the coastline and cont. slope
cont <- st_as_sf(as.contour(bathy, ))
coast <- cont %>% filter(level == 0)
cont_slope <- cont %>% filter(level == -2000) #Is 2000m the correct depth for continental shelf?
cont_slope_rast <- rasterize(cont_slope, agg, field = "level")
plot(agg)
plot(cont_slope_rast, col = "red")
cont_slope_rast <- rasterize(cont_slope, agg)
plot(cont_slope_rast, col = "red")

ggplot() + 
  geom_sf(data = cont, col="black") + 
  geom_sf(data = coast, col = "red") +
  geom_sf(data = cont_slope, col = "purple") +
  #geom_sf(data = whales_8_10sf, mapping = aes()) +
  scale_color_gradient(low="green", high="red")
```

>>>>

print(ggplot(data = pivot1, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 3, color = "black") +
  theme_classic() +
  ggtitle(paste0("Distribution of third octave level intensities: Drift ", substr(ft,10,11))) +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.7,0.8)) +
  (facet_grid(~factor(SL, levels = c('BB_20.24000', 'TOL_63', 'TOL_125', 'TOL_2000', 'TOL_5000', 'TOL_20000')))))
}

@@

#SL$size_f <- factor(SL$ , levels = c('BB_20.2400', 'TOL_63', 'TOL_125', 'TOL_2000', 'TOL_5000', 'TOL_20000'))))

ggplot(data = pivot3, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 3, color = "black") +
  theme_classic() +
  ggtitle("Distribution of third octave level intensities: All Drifts Combined") +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.7,0.8)) +
  facet_grid(~factor(SL, levels = c('BB_20.24000', 'TOL_63', 'TOL_125', 'TOL_2000', 'TOL_5000', 'TOL_20000')))

@@

Avg sound levels by band (scatterplot)
```{r}
avgdB <- TO_23 %>%
  summarize(TOL_25 = mean(TOL_25),
            TOL_31.5 = mean(TOL_31.5),
            TOL_40 = mean(TOL_40),
            TOL_50 = mean(TOL_50),
            TOL_63 = mean(TOL_63),
            TOL_80 = mean(TOL_80),
            TOL_100 = mean(TOL_100),
            TOL_125 = mean(TOL_125),
            TOL_160 = mean(TOL_160),
            TOL_200 = mean(TOL_200),
            TOL_250 = mean(TOL_250),
            TOL_315 = mean(TOL_315),
            TOL_400 = mean(TOL_400),
            TOL_500 = mean(TOL_500),
            TOL_630 = mean(TOL_630),
            TOL_800 = mean(TOL_800),
            TOL_1000 = mean(TOL_1000),
            TOL_1250 = mean(TOL_1250),
            TOL_1600 = mean(TOL_1600),
            TOL_2000 = mean(TOL_2000),
            TOL_2500 = mean(TOL_2500),
            TOL_3150 = mean(TOL_3150),
            TOL_4000 = mean(TOL_4000),
            TOL_5000 = mean(TOL_5000),
            TOL_6300 = mean(TOL_6300),
            TOL_8000 = mean(TOL_8000),
            TOL_10000 = mean(TOL_10000),
            TOL_12500 = mean(TOL_12500),
            TOL_16000 = mean(TOL_16000),
            TOL_20000 = mean(TOL_20000)) %>%
  pivot_longer(cols = TOL_25:TOL_20000,
               names_to = "frequency",
               values_to="dB")

ggplot(avgdB, aes(x = factor(frequency, levels = unique(frequency)), y = dB)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Box plots
```{r}
TO_longer <- pivot_longer(TO_10, cols = TOL_25:TOL_20000,
               names_to = "frequency",
               values_to="dB")
ggplot(TO_longer, aes(x = factor(frequency, levels = unique(frequency)), y = dB)) + 
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Violin plot
ggplot(TO_longer, aes(x = factor(frequency, levels = unique(frequency)), y = dB)) + 
  geom_violin() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

@@

```{r}
model3 <- gam(BWpresence ~ depth + TOL_125 + dist2shore + mldTemp, 
              data = allDrifts, family = binomial)
#plot(model3)

@@
